Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2013-03-04 doc/paper-lcr.tex
fixed conflict

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight. In
time delay gravitational lenses, ``external convergence,'' $\kappax$, can
dominate the uncertainty in the inferred distances and hence cosmological parameters.
 In this paper we attempt
to reconstruct $\kappax$, due to line of sight structure, using a simple halo
model.
%
We use mock catalogues from the Millennium Simulation, and calibrate and 
compare our reconstructed $\pr(\kappax)$ to ray-traced $\kappax$ ``truth''
values; taking into account realistic uncertainties on redshift and stellar
masses. \tom{We find that the reconstruction of $\kappax$ provides an improvement in
precision of $\sim$50\% over galaxy number counts.} 
%
We find that the lowest-$\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$. In anticipation of future samples with thousands of lenses,
we find that selecting the third of the systems with the
highest precision $\kappax$ estimates gives a  sub-sample of unbiased time
delay distance measurements \phil{with (on average) just} 1\% uncertainty due
to line of sight external convergence effects. Photometric data alone are
sufficient to pre-select the best-constrained lines of sight, and can be done
before investment in lightcurve monitoring. 
%
\phil{Conversely, we show that selecting lines of sight
with high external shear could, with the reconstruction model presented here,
induce biases of up to 1\% in time delay distance.} \new{We find that a major potential source of systematic error is uncertainty in the high mass end of the stellar mass-halo mass relation; this could introduce ~2\% biases on the time-delay distance if completely ignored.} We suggest areas for the
improvement of this general analysis framework (including more sophisticated
treatment of high mass structures) that should allow yet more accurate
cosmological inferences to be made.

@@

To estimate the potential size of this systematic error, we repeat the
analysis of the previous section, still using the  the
\citet{BehrooziEtal2010} relation to generate the stellar masses of our
calibration lines of sight and hence carry out the inference, but now using
simulated catalog datasets that were generated from the stellar mass--halo
mass relation of \citet{MosterEtal2010}.

@@

The systematic errors reported here are perhaps overly pessimistic, because
the bias is due to differences in the stellar masses predicted for high mass
halos. If new observations can discriminate between the high mass end of the
\citet{BehrooziEtal2010} and \citet{MosterEtal2010} relations, then the
systematic error on the reconstruction will decrease. Similarly, incorporating
additional information about the high mass group and cluster-scale halo
systems, such as their richness or other occupation statistics, would help
mitigate the stellar to halo mass conversion error. }

@@

\item \new{Reconstructing a line of sight with an incorrect stellar mass--halo mass relation can introduce systematic errors on the inferrred time-delay distance at the percent level. This error can only be reduced by tightening the observational limits on the stellar mass--halo mass relation at the high mass end.}

>>>>

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight. In
time delay gravitational lenses, ``external convergence,'' $\kappax$, can
dominate the uncertainty in the inferred distances and hence cosmological parameters.
 In this paper we attempt
to reconstruct $\kappax$, due to line of sight structure, using a simple halo
model.
%
We use mock catalogues from the Millennium Simulation, and calibrate and 
compare our reconstructed $\pr(\kappax)$ to ray-traced $\kappax$ ``truth''
values; taking into account realistic uncertainties on redshift and stellar
masses. \tom{We find that the reconstruction of $\kappax$ provides an improvement in
precision of $\sim$50\% over galaxy number counts.} 
%
We find that the lowest-$\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$. In anticipation of future samples with thousands of lenses,
we find that selecting the third of the systems with the
highest precision $\kappax$ estimates gives a  sub-sample of unbiased time
delay distance measurements \phil{with (on average) just} 1\% uncertainty due
to line of sight external convergence effects. Photometric data alone are
sufficient to pre-select the best-constrained lines of sight, and can be done
before investment in lightcurve monitoring. 
%
\phil{Conversely, we show that selecting lines of sight
with high external shear could, with the reconstruction model presented here,
induce biases of up to 1\% in time delay distance.} \new{We find that a major potential source of systematic error is uncertainty in the high mass end of the stellar mass-halo mass relation; this could introduce $\sim$2\% biases on the time-delay distance if completely ignored.} We suggest areas for the
improvement of this general analysis framework (including more sophisticated
treatment of high mass structures) that should allow yet more accurate
cosmological inferences to be made.

@@

To estimate the potential size of this systematic error, we repeat the
analysis of the previous section, still using the  the
Behroozi et al. relation to generate the stellar masses of our
calibration lines of sight and hence carry out the inference, but now using
simulated catalog datasets that were generated from the stellar mass--halo
mass relation of Moster et al.

@@

The systematic errors reported here are perhaps overly pessimistic, because
the bias is due to differences in the stellar masses predicted for high mass
halos. If new observations can discriminate between the high mass end of the
Behroozi et al. and Moster et al. relations, then the
systematic error on the reconstruction will decrease. Similarly, incorporating
additional information about the high mass group and cluster-scale halo
systems, such as their richness or other occupation statistics, would help
mitigate the stellar to halo mass conversion error. }

@@

\item \new{Reconstructing a line of sight with an incorrect stellar mass--halo mass relation can introduce systematic errors on the inferrred time-delay distance at the percent level. This error can be reduced by tightening the observational limits on the stellar mass--halo mass relation at the high mass end.}
